---
title: "Data analysis DDR-OSR"
author: "Elsje Burgers"
last modified: "2024-11-26"
output: html_document
---

This script can be used to obtain the data used for figures and parameter fitting.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}
rm(list = ls())
library(tidyverse)

save_data = 0 #if 1 data will be saved as csv, if 0 data will not be saved
```

# Load data
Here, we load the files and correct some minor mistakes in the files to make sure they all use the same way to indicate the cell line, reporter etc.
```{r}
data_GFP = do.call(rbind, lapply(list.files(path = "..//Data//all_data//GFP//", 
                                            full.names = TRUE), read.delim))
data_ANVPI = do.call(rbind, lapply(list.files(path = "..//Data//all_data//AnVPI//", 
                                              full.names = TRUE), read.delim))

data_GFP$cell_line[data_GFP$cell_line == "HepG2MDM2-GFP"] = "HepG2 MDM2-GFP"
data_GFP$cell_line[data_GFP$cell_line == "HepG2p21-GFP"] = "HepG2 p21-GFP"
data_GFP$cell_line[data_GFP$cell_line == "HepG2 NRF-2-GFP"] = "HepG2 NRF2-GFP"

data_ANVPI$cell_line[data_ANVPI$cell_line == "HepG2MDM2-GFP"] = "HepG2 MDM2-GFP"
data_ANVPI$cell_line[data_ANVPI$cell_line == "HepG2p21-GFP"] = "HepG2 p21-GFP"
data_ANVPI$cell_line[data_ANVPI$cell_line == "HepG2 NRF-2-GFP"] = "HepG2 NRF2-GFP"
data_ANVPI$replID[data_ANVPI$plateID == "20180329_SBT-TQ_ATF4_rep2"] = 2

data_GFP = mutate(data_GFP, treatment = gsub(' ', '', treatment))
data_ANVPI = mutate(data_ANVPI, treatment = gsub(' ', '', treatment))
```

# GFP data
## Correct and normalise
The data is corrected for the background GFP expression by subtracting the intensity for the exposure to 0.01 DMSO from the measured intensity.
There is various ways of normalising the data. In the following chunk it is normalised using the minimal value per plate and the mean value per plate.
```{r message=FALSE, warning=FALSE}
# Data set containing the 0.01 DMSO values is defined here.
correction = filter(data_GFP, treatment=="DMSO" & dose_uM==0.01) %>%
  select(c(3:7,9)) %>%
  rename(DMSO_value = GFP_int)

# Now the correction is done.
GFP_cor = data_GFP %>%
  left_join(correction) %>%
  mutate(GFP_corrected = ifelse(!is.na(GFP_int),
                          GFP_int - DMSO_value,
                          GFP_int))

# Here the mean and min value to normalise with are extracted.
plate_mean_min <- GFP_cor %>%
  group_by(cell_line, replID) %>%
  summarise(mean_value = mean(GFP_corrected, na.rm = TRUE),
            min_value = min(GFP_corrected, na.rm = TRUE))

# Min-mean normalization
GFP_normalised <- GFP_cor %>%
  left_join(plate_mean_min) %>%
  mutate(GFP_norm = ifelse(!is.na(GFP_corrected),
                           (GFP_corrected - min_value)/(mean_value - min_value),
                           GFP_corrected))

# Get desired dataset
GFP_realdata = GFP_normalised %>%
  mutate(StateVar = gsub('HepG2 ', '', cell_line )) %>%
  mutate(StateVar = gsub('-GFP', '', StateVar )) %>%
  select(c(1, 2, 5, 6, 7, 16, 17)) %>%
  rename(data4modelReal = GFP_norm, timeID_orig = timeID) %>%
  mutate(timepoints = round(timeAfterExposure))
```

## Interpolate
We interpolate the data using a cubic normal spline.
```{r}
tmps = GFP_realdata %>%
  group_by(StateVar, replID) %>%
  summarise(last_tmps = max(timeAfterExposure), first_tmps = min(timeAfterExposure))
max_tmp = min(tmps$last_tmps)

# Now we interpolate the data
interpolation  = GFP_realdata %>%
  group_by(treatment, dose_uM, replID,  StateVar) %>%
  summarise(timepoints = spline(x = timeAfterExposure, y = data4modelReal, 
                                xout = 0:max_tmp, method = "natural")$x, 
            data4modelInterpol = spline(x = timeAfterExposure, y = data4modelReal, 
            xout = 0:max_tmp, method = "natural")$y) %>%
  left_join(GFP_realdata) %>%
  group_by(treatment, dose_uM, replID,  StateVar) %>%
  mutate(data4modelInterpol = ifelse(timepoints < min(timeAfterExposure, na.rm = TRUE)
                                     & is.na(timeAfterExposure),NA, data4modelInterpol))
```
## Filter
We select the compound and doses of interest.
```{r}
compound = "ETO" #"DEM"
max_dose  = 25 #200

selected_data = interpolation %>%
    filter(treatment == compound & dose_uM <= max_dose)
```

## Plot
We plot the data to check if everything went as planned.
```{r}
# plot all timepoints
plot = ggplot(data = selected_data) + 
  geom_point(aes(x = timepoints, y = data4modelInterpol, color  = as.factor(replID)), shape = 1) + 
  geom_point(aes(x = timeAfterExposure, y = data4modelReal, color  = as.factor(replID)), shape = 2) +
  facet_grid(cols = vars(dose_uM), rows = vars(StateVar), scales = "free")
print(plot)

# plot early timepoints, most errors in extrapolation occur there.
plot = ggplot(data = filter(selected_data, timepoints<10)) + 
  geom_point(aes(x = timepoints, y = data4modelInterpol, color  = as.factor(replID)), shape = 1) + 
  geom_point(aes(x = timeAfterExposure, y = data4modelReal, color  = as.factor(replID)), shape = 3) +
  facet_grid(cols = vars(dose_uM), rows = vars(StateVar), scales = "free")
print(plot)

# for one protein
plot = ggplot(data = filter(selected_data, timepoints<20 & StateVar == "NRF2" )) + 
  geom_point(aes(x = timepoints, y = data4modelInterpol, 
                 shape  = as.factor(replID)), color = "black") +
  geom_point(aes(x = timeAfterExposure, y = data4modelReal, 
                 shape  = as.factor(replID)), color = "red") +
  labs(shape = "replID") + 
  facet_wrap(vars(dose_uM))
print(plot)
```
## Save
If desired, we can save the data here in the format required for the parameter fitting.
```{r}
fitdata = selected_data %>%
  mutate(timeID = timepoints+1, dose_uMadj = dose_uM) %>%
  select("StateVar", "replID", "treatment", "timeID", "timeAfterExposure", 
         "dose_uM", "dose_uMadj", "timepoints", "data4modelInterpol", 
         "data4modelReal", "timeID_orig")

if (save_data == 1){
  write.csv(fitdata, file = paste("..\\Data\\GFPdata_", compound, "_", 
                                  Sys.Date(), ".csv", sep = ""))
}
```

# Viability data
This data does not need to be corrected or interpolated. However, to make it easier to work with, we do want to filter and average the data. Also, we will normalise the cell count.
## Filter
```{r}
data_ANVPI_filter = data_ANVPI %>%
  filter(treatment == compound & timeID<43)
```

## Normalise
```{r}
count0 = data_ANVPI_filter %>%
  filter(timeID == 1) %>%
  rename(count_t0 = cell_count) %>%
  select(treatment, dose_uM, replID, cell_line, count_t0)

data_ANVPI_norm = data_ANVPI_filter %>%
  left_join(count0) %>%
  mutate(count_norm = cell_count/count_t0)
```

## Average
```{r}
data_ANVPI_mean = data_ANVPI_norm %>%
  group_by(treatment, dose_uM, timeID) %>%
  summarise(mean_time = mean(timeAfterExposure, na.rm = TRUE),
            mean_count = mean(cell_count, na.rm = TRUE),
            sd_count = sd(cell_count, na.rm = TRUE),
            mean_count_norm = mean(count_norm, na.rm = TRUE),
            sd_count_norm = sd(count_norm, na.rm = TRUE),
            mean_PI = mean(PIPos_0.1, na.rm = TRUE),
            sd_PI = sd(PIPos_0.1, na.rm = TRUE),
            mean_AnV = mean(AnVPos_0.1, na.rm = TRUE),
            sd_AnV = sd(AnVPos_0.1, na.rm = TRUE))
```
## Save
If desired, we can save the data here. 
```{r}
if (save_data == 1){
  write.csv(data_ANVPI_mean, file = paste("..\\Data\\ViabilityData_", compound, 
                                          "_", Sys.Date(), ".csv", sep = ""))
}
```

```{r}
sessionInfo()
```
R version 4.3.3 (2024-02-29 ucrt)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 11 x64 (build 22631)

Matrix products: default


locale:
[1] LC_COLLATE=English_United Kingdom.utf8  LC_CTYPE=English_United Kingdom.utf8   
[3] LC_MONETARY=English_United Kingdom.utf8 LC_NUMERIC=C                           
[5] LC_TIME=English_United Kingdom.utf8    

time zone: Europe/Amsterdam
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lubridate_1.9.3 forcats_1.0.0   stringr_1.5.1   dplyr_1.1.4     purrr_1.0.2     readr_2.1.5    
 [7] tidyr_1.3.1     tibble_3.2.1    ggplot2_3.5.1   tidyverse_2.0.0

loaded via a namespace (and not attached):
 [1] crayon_1.5.3      vctrs_0.6.5       cli_3.6.3         knitr_1.49        rlang_1.1.4       xfun_0.49        
 [7] stringi_1.8.4     generics_0.1.3    labeling_0.4.3    glue_1.8.0        colorspace_2.1-1  hms_1.1.3        
[13] scales_1.3.0      fansi_1.0.6       grid_4.3.3        munsell_0.5.1     evaluate_1.0.1    tzdb_0.4.0       
[19] lifecycle_1.0.4   compiler_4.3.3    timechange_0.3.0  pkgconfig_2.0.3   rstudioapi_0.17.1 farver_2.1.2     
[25] R6_2.5.1          tidyselect_1.2.1  utf8_1.2.4        pillar_1.9.0      magrittr_2.0.3    tools_4.3.3      
[31] withr_3.0.2       gtable_0.3.6
